---
title: "Australian and Canadian Greenhouse Gas Emissions"

output:
  html_document:
    df_print: paged
    fig_width: 7
    fig_height: 5

editor_options: 
  chunk_output_type: inline

date: 2018-12-30T00:00:00
lastmod: 2019-01-09T00:00:00

slug: "EmissionSources"

draft: true 

---

```{r setup,include=FALSE}
knitr::opts_chunk$set(echo=FALSE,include=FALSE)
#knitr::opts_knit$set(root.dir = normalizePath(".."))

library(readr)
library(tidyr)
library(dplyr)
library(reshape2)
library(readxl)
library(ggrepel)
library(knitr)
library("RColorBrewer")
library(ggplot2)
library(stringr) 
rm(list=ls())
pal <- function(col, border = "light gray", ...){
  n <- length(col)
  plot(0, 0, type="n", xlim = c(0, 1), ylim = c(0, 1),
       axes = FALSE, xlab = "", ylab = "", ...)
  rect(0:(n-1)/n, 0, 1:n/n, 1, col = col, border = border)
}
```

The conversation around greenhouse gas emissions in Australia differs from that in Canada. In Australia there is much attention on coal generation and renewable energy, where as in Canada the focus is on oil and gas, and in particular the oil sands and export pipelines. To get a sense of where emissions come from in each country this post breaks down and compares emission sources.

```{r}
##########################################
## Import Canadian IPCC Data
##########################################
# Import sector definitions crosswalk
getwd()
CDN_IPCC_def <- read_excel("Data_AUCDN_GHGBreakdown/CanadaAU_SectorCrosswalk.xlsx",sheet = "CDN_IPCC",range = "A3:D85")

# Import 2018 CDN NIR IPCC data
CDN_IPCC_Data <- read_excel("Data_AUCDN_GHGBreakdown/2018NIR_IPCC_Tables_CA_1990-2016.xlsx",sheet = "Summary",range = "A5:AF89")
colnames(CDN_IPCC_Data)[1] <- "CDN_Cat1"
colnames(CDN_IPCC_Data)[2] <- "CDN_Cat2"
colnames(CDN_IPCC_Data)[3] <- "CDN_Cat3"
CDN_IPCC_Data <- select(CDN_IPCC_Data,-X__3,-X__4)

#Merge in IPCC definitions
CDN_IPCC_Data <- merge(CDN_IPCC_Data,CDN_IPCC_def,by=c("CDN_Cat1","CDN_Cat2","CDN_Cat3"),all=FALSE)
CDN_IPCC_Data <- rename(CDN_IPCC_Data,"IPCC_Sector"="IPCC NIR Sector")

#Merge in useful definitions to use in place of IPCC sectors
Sector_def <- read_excel("Data_AUCDN_GHGBreakdown/CanadaAU_SectorCrosswalk.xlsx",sheet = "List of sectors")
Sector_def <- rename(Sector_def,"IPCC_Sector"="IPCC NIR Sector")
Sector_def <- filter(Sector_def,IPCC_Sector!="From IPCC Sector Definitions for NIR CIF Tables")

#use Convenient_Name2 renamed as Convenient_Name
CDN_IPCC_Data <- merge(CDN_IPCC_Data,Sector_def,by=c("IPCC_Sector"),all=FALSE)
CDN_IPCC_Data <- select(CDN_IPCC_Data,-IPCC_Sector,-CDN_Cat1,-CDN_Cat2,-CDN_Cat3,-Convenient_Name)
CDN_IPCC_Data <- rename(CDN_IPCC_Data,Convenient_Name=Convenient_Name2)

#reshape into long format
CDN_IPCC_Data <- CDN_IPCC_Data %>% gather("year","GHGs",-GHGCat,-SubCat1,-SubCat2,-SubCat3,-Convenient_Name) %>% transform(GHGs = as.numeric(GHGs))  %>% transform(year = as.numeric(year))

#Convert GHGs to MtCO2eq from ktCO2eq
CDN_IPCC_Data$GHGs <- CDN_IPCC_Data$GHGs/1000

#drop values with missing GHGs
CDN_IPCC_Data <- filter(CDN_IPCC_Data,!is.na(CDN_IPCC_Data$GHGs))

#replace all is.na with "Total" so that I can use these as grouping variables in aggregate
CDN_IPCC_Data[is.na(CDN_IPCC_Data)] <- "Total"
#str(CDN_IPCC_Data)

#aggregate building category 
CDN_IPCC_Data$SubCat2[CDN_IPCC_Data$SubCat2=="Commercial and Institutional"] <- "Buildings" 
CDN_IPCC_Data$SubCat2[CDN_IPCC_Data$SubCat2=="Residential"] <- "Buildings"
CDN_IPCC_Data <- aggregate(GHGs~.,data=CDN_IPCC_Data,FUN=sum)

#change GHGCat text to lowercase
CDN_IPCC_Data$GHGCat <- sapply(CDN_IPCC_Data$GHGCat,FUN=tolower)
library(Hmisc)
CDN_IPCC_Data$GHGCat <- capitalize(CDN_IPCC_Data$GHGCat)

#This data can be used for both country-sector break downs, and projections. But for breakdowns (future post) might want to use 2018 NIR data to get 1 more year.

```

```{r,message=FALSE,warnings=FALSE}
#AU Data to compare to Canadian, using same color schemes
#Import AU sector def from the crosswalk
AU_SectorDef <- read_excel("Data_AUCDN_GHGBreakdown/CanadaAU_SectorCrosswalk.xlsx",sheet = "AU_IPCC",range = "A2:G91")

#Import all AU AGEIS data, and assign categories from AU_SectorDef by row number. This is because the AU data contains "Other" multiple times. 
AU_total <- NULL
for (YearName in c(1990:2016)) {
  print(paste0("AU",YearName))
  temp <- read_csv(paste0("Data_AUCDN_GHGBreakdown/AGEISbyYear/AU",YearName,".csv"),col_names = TRUE)[ ,1:2]
  temp <- temp %>% filter(is.na(temp[2])!=1)
  temp[,3] <- YearName
  temp[,4] <- seq.int(nrow(temp))
  AU_total <- rbind(AU_total,temp)
}

colnames(AU_total)[2] <- "GHGs"
AU_total <- rename(AU_total,"year"="V3")
AU_total <- rename(AU_total,"ColumnID"="V4")

AU_total$GHGs[AU_total$GHGs=="Data is not available"] <- NA
AU_total$GHGs[AU_total$GHGs=="Data is confidential"] <- NA
str(AU_total)
AU_total$GHGs <- as.numeric(gsub(",","",AU_total$GHGs))
#Convert to numeric
AU_total <- AU_total %>% transform(GHGs = as.numeric(GHGs))  %>% transform(year = as.numeric(year))

#convert to million tonnes (Mt) of CO2
AU_total$GHGs <- AU_total$GHGs*1000 /(10^6)

#Merge IPCC sector definitions from AU_SectorDef
AU_total <- merge(AU_total,AU_SectorDef,by=c("ColumnID","Category"))
AU_total <- rename(AU_total,"IPCC_Sector"="IPCC NIR Sector")
AU_total <- select(AU_total,year,GHGs,IPCC_Sector)

#Merge in convenient sector defintions
AU_IPCC_Data <- merge(AU_total,Sector_def,by=c("IPCC_Sector"),all=FALSE)
AU_IPCC_Data <- select(AU_IPCC_Data,-IPCC_Sector,-Convenient_Name)
AU_IPCC_Data <- rename(AU_IPCC_Data,Convenient_Name=Convenient_Name2)

#change GHGCat text to lowercase
AU_IPCC_Data$GHGCat <- sapply(AU_IPCC_Data$GHGCat,FUN=tolower)
library(Hmisc)
AU_IPCC_Data$GHGCat <- capitalize(AU_IPCC_Data$GHGCat)

#drop missing GHG data. This keeps GHGs as a numerical variable after replacing NA with "Total"
AU_IPCC_Data <- filter(AU_IPCC_Data,!is.na(AU_IPCC_Data$GHGs))

#replace all is.na with "Total" so that I can use these as grouping variables in aggregate
AU_IPCC_Data[is.na(AU_IPCC_Data)] <- "Total"
#str(AU_IPCC_Data)

#aggregate building category 
AU_IPCC_Data$SubCat2[AU_IPCC_Data$SubCat2=="Commercial and Institutional"] <- "Buildings" 
AU_IPCC_Data$SubCat2[AU_IPCC_Data$SubCat2=="Residential"] <- "Buildings"
AU_IPCC_Data <- aggregate(GHGs~.,data=AU_IPCC_Data,FUN=sum)

#complete AU_IPCC_Data ready for plottings

```


```{r}
#Combine AU and Canadian Data
temp1 <- mutate(AU_IPCC_Data,cty="AU")
temp2 <- mutate(CDN_IPCC_Data,cty="CDN")
IPCC_Data <- rbind(temp1,temp2)
rm(temp1,temp2)

#rename GHGcat
IPCC_Data$GHGCat[IPCC_Data$GHGCat=="Industrial processes and product use"] <- "Industrial processes"
IPCC_Data$GHGCat[IPCC_Data$GHGCat=="Land use, land-use change and forestry"] <- "LULUCF"
```

Here's Australia's and Canada's emissions since 1990. The importance of energy emissions, and the growth in them since 1990, is clear. *Energy* is primarily the burning of fossil fuels for electricity, heat, and transportation. *Agriculture* includes emissions like methane, but not energy used in farm tractors, to heat greenhouses, etc. *Industrial proceses* are *Waste* emissions come directly from sources like cement or landfills, but like agriculture do not include associated fossil fuel use. *LULUCF* is land use, land-use change, and forestry. 

Data for Canada comes from the [2018 National Inventory Report](http://data.ec.gc.ca/data/substances/monitor/national-and-provincial-territorial-greenhouse-gas-emission-tables/), and for Australia from the [National Greenhouse Gas Inventory via AGEIS](http://ageis.climatechange.gov.au/NGGI.aspx). 


```{r,include=TRUE}
# Canada and AU 5 categories of emissions
IPCC_Dataplot <- IPCC_Data %>% filter(GHGCat!="Total" & SubCat1=="Total") %>% select(GHGCat,year,GHGs,cty)

#Sort by 2016 size: need order of GHGCat by the 2016 size
factororder <- filter(IPCC_Dataplot,IPCC_Dataplot$year==2016 & cty=="AU") %>% arrange(-GHGs)
factororder <- as.character(factororder$GHGCat)

ggplot(IPCC_Dataplot,aes(year,GHGs,fill=factor(IPCC_Dataplot$GHGCat,levels=factororder,ordered=TRUE))) + geom_area() + theme_bw() + scale_fill_brewer(palette="Spectral") + 
  labs(title="Greenhouse gas emissions", subtitle=element_blank(), 
       y=expression('Emissions (Mt CO'[2]*'eq)'),
       x=element_blank(),
       fill="Category") +
  facet_grid(cols=vars(cty),labeller=as_labeller(c(AU="Australia",CDN="Canadian"))) 

```


```{r,include=FALSE}
#Plot by second set of major categories within energy
IPCC_Dataplot <- IPCC_Data %>% filter(GHGCat=="Energy" & SubCat1!="Total" & SubCat2=="Total") %>% select(GHGCat,SubCat1, SubCat2,year,GHGs,cty)

factororder <- filter(IPCC_Dataplot,IPCC_Dataplot$year==2016 & cty=="AU") %>% arrange(-GHGs)
factororder <- as.character(factororder$SubCat1)

ggplot(IPCC_Dataplot,aes(year,GHGs,fill=factor(IPCC_Dataplot$SubCat1,levels=factororder,ordered=TRUE))) + geom_area() + theme_bw() + scale_fill_brewer(palette="Spectral") + 
  labs(title="1990-2016 greenhouse gas emissions", subtitle="Totals Within Energy Category", 
       y=expression('Emissions (MT CO'[2]*')'),
       x=element_blank(),
       fill="Industry") +
  facet_grid(cols=vars(cty),labeller=as_labeller(c(AU="Australia",CDN="Canadian"))) 
```

Emissions within the *Energy* category come from a variety of sources. I've plotted these emissions below, ordered largest to smallest by Australia's 2016 emissions. Electricity and other fossil fuel related emissions are red-orange, and transportation emissions are in blue. 

In Australia, electricity generation is the largest source of emissions owing to the widespread use of coal. The importance of coal industry is further evidence in fugitive emissions from coal mining being Australia's 4th largest emissions source. Canada, with its relatively clean electricity emissions from legacy hydro power, has road transport as the largest source. While fugitive coal emissions are negligible in Canada, the importance of oil and gas can be seen through the large fossil fuel extraction and fugitive oil and gas emissions. 

```{r,include=TRUE}
#subset of data for plotting
IPCC_Dataplot <- IPCC_Data %>% filter(GHGCat=="Energy" & SubCat1!="Total") %>% filter((SubCat2!="Total" & SubCat2!="Energy Industries" & SubCat3=="Total") | (SubCat2=="Energy Industries" & SubCat3!="Total"))

#Sort by 2016 size: need order of GHGCat by the 2016 size
factororder <- filter(IPCC_Dataplot,IPCC_Dataplot$year==2016 & cty=="AU") %>% arrange(-GHGs)
factororder <- as.character(factororder$Convenient_Name)

#Generate sets of colors using colorRampePalette, append to the factors
transport_colors=c("#2171B5","#7EB3D6")
transport_colors = colorRampPalette(transport_colors)(5)
#pal(transport_colors)

FF_colors=c("#DC050C","#F6C141")
FF_colors = colorRampPalette(FF_colors)(5)
#pal(FF_colors)

factorcolors <- append(c("#4EB265","#B178A6","#D6C1DE"),append(FF_colors,transport_colors))
#pal(factorcolors)

#assign names from factororder to factorcolors
names(factorcolors) <- c(factororder[9],factororder[3],factororder[7],factororder[1],factororder[4],factororder[5],factororder[6],factororder[11],factororder[2],factororder[8],factororder[10],factororder[12],factororder[13])

ggplot(IPCC_Dataplot,aes(year,GHGs,fill=factor(IPCC_Dataplot$Convenient_Name,levels=factororder,ordered=TRUE))) + geom_area() + theme_bw() + scale_fill_manual(values=factorcolors) + 
  labs(title="Greenhouse gas emissions", subtitle="Within Energy emissions", 
       y=expression('Emissions (MT CO'[2]*'eq)'),
       x=element_blank(),
       fill="Subcategory") +
  facet_grid(cols=vars(cty),labeller=as_labeller(c(AU="Australia",CDN="Canadian"))) 
```

#### Provincial & State emissions

Emission sources differ substantially across Australia and Canada. Below, I plot all emissions excluding LULUCF by province/state, emission category, and year. I've separated *Energy* into electricity, transport, fossil fuel extraction activities, buildings, and manufacturing sources. For Canada, Alberta jumps out for both its high emissions and large share from the oil and gas industry. The importance of hydro power to BC and Quebec is clear in the lack of emissions from electricity generation. Emissions across Australian states are relatively uniform compared to Canada, though with South Australia and Western Australia having a smaller share from electricity generation. 

```{r}
## Import Canadian province data and merge in IPCC definitions
ProvinceList <- c("NL","PE","NS","NB","QC","ON","MB","SK","AB","BC")
#ProvinceList <- c("AB","BC")
CDNProvGHGs <- NULL
for (ProvName in ProvinceList) {
  filename <- paste0("2018NIR_Tables_",ProvName,"_1990-2016_Unrounded.xlsx")
  print(filename)
  temp <- read_excel(paste0("Data_AUCDN_GHGBreakdown/C-Tables-IPCC-Sector-Provinces-Territories/",filename),sheet="Summary",skip = 4)
  #process the imported xlsx file before appending to the CDNProvGHGs
  colnames(temp)[1] <- "CDN_Cat1"
  colnames(temp)[2] <- "CDN_Cat2"
  colnames(temp)[3] <- "CDN_Cat3"
  #format temp data for rbind
  temp <- temp %>% select(-X__3,-X__4) %>% mutate(region=ProvName)
  temp <- filter(temp,!is.na(temp$"2016"))
  CDNProvGHGs <- rbind(CDNProvGHGs,temp)  
}  

#Merge in IPCC definitions. This requires adjusting a few categories so that the names from the 2018NIR_Tables matches the IPCC crosswalk
CDN_IPCC_def$CDN_Cat1[CDN_IPCC_def$CDN_Cat1=="TOTAL1"] <- "TOTAL"
CDN_IPCC_def$CDN_Cat2[CDN_IPCC_def$CDN_Cat2=="Transport2"] <- "Transport1"
CDN_IPCC_def$CDN_Cat2[CDN_IPCC_def$CDN_Cat2=="Chemical Industry"] <- "Chemical Industry2"

CDNProvGHGs <- merge(CDNProvGHGs,CDN_IPCC_def,by=c("CDN_Cat1","CDN_Cat2","CDN_Cat3"),all=FALSE)
CDNProvGHGs <- rename(CDNProvGHGs,"IPCC_Sector"="IPCC NIR Sector")

CDNProvGHGs <- select(CDNProvGHGs,-CDN_Cat1,-CDN_Cat2,-CDN_Cat3)
CDNProvGHGs <- mutate(CDNProvGHGs,cty="CDN")

#gather data
CDNProvGHGs <- CDNProvGHGs %>% gather("year","GHGs",-IPCC_Sector,-region,-cty)

#filter out data not used - e.g., that without an IPCC sector
CDNProvGHGs <- filter(CDNProvGHGs,!(CDNProvGHGs$GHGs=="-" | CDNProvGHGs$GHGs=="x") & !is.na(CDNProvGHGs$IPCC_Sector))
CDNProvGHGs$year <- as.numeric(CDNProvGHGs$year)
CDNProvGHGs$GHGs <- as.numeric(CDNProvGHGs$GHGs)

#unique(CDNProvGHGs$IPCC_Sector)
#create full set of categories
#simplify dataset:
CDNProvGHGs <- filter(CDNProvGHGs,IPCC_Sector %in% c("Total","1.A.1.a Public Electricity and Heat Production","1.A.2 Manufacturing Industries and Construction","1.A.3 Transport","1.A.4.c Agriculture, Forestry and Fishing","1.A.4.b Residential","1.A.4.a Commercial / Institutional","1.B Fugitive Emissions From Fuels","2. Industrial Processes","3. Agriculture","5. Waste","1.A.1.b Petroleum Refining","1.A.1.c Manufacture of Solid Fuels and Other Energy Industries"))

CDNProvGHGs$IPCC_Sector[CDNProvGHGs$IPCC_Sector=="1.A.1.b Petroleum Refining"] <- "1.A.1.z Fossil Fuel Extraction and Refining"
CDNProvGHGs$IPCC_Sector[CDNProvGHGs$IPCC_Sector=="1.A.1.c Manufacture of Solid Fuels and Other Energy Industries"] <- "1.A.1.z Fossil Fuel Extraction and Refining"

CDNProvGHGs$IPCC_Sector[CDNProvGHGs$IPCC_Sector=="1.A.4.a Commercial / Institutional" ] <- "1.A.4 Buildings"
CDNProvGHGs$IPCC_Sector[CDNProvGHGs$IPCC_Sector=="1.A.4.b Residential"] <- "1.A.4 Buildings"
CDNProvGHGs$IPCC_Sector[CDNProvGHGs$IPCC_Sector=="1.A.4.c Agriculture, Forestry and Fishing"] <- "1.A.4 Buildings"

#aggregate over redefined categories
CDNProvGHGs <- aggregate(GHGs~.,data=CDNProvGHGs,FUN=sum)

#verify totals add close to total within each region-year? 
temp <- subset(CDNProvGHGs,CDNProvGHGs$IPCC_Sector=="Total")
tempsum <- subset(CDNProvGHGs,CDNProvGHGs$IPCC_Sector!="Total") %>% select(-cty) %>% group_by(region,year) %>% summarise(sectorsum=sum(GHGs))

ggplot() + theme_bw() + facet_grid(cols=vars(region)) +
  geom_line(data=temp,aes(year,GHGs),color='red',size=5) +
  geom_line(data=tempsum,aes(year,sectorsum),color='green')
#totals match up as expected!!
# "Total" (to verify totals)
# "1.A.1.a Public Electricity and Heat Production"
# "1.A.1.z Fossil Fuel Extraction" (to create this!)
#   from sum of:
#   1.A.1.b Petroleum Refining
#   1.A.1.c Manufacture of Solid Fuels and Other Energy Industries  
# "1.A.2 Manufacturing Industries and Construction" 
# "1.A.3 Transport" 
# 1.A.4 Buildings (created from sum of:)
#   "1.A.4.c Agriculture, Forestry and Fishing"
                                           
#   "1.A.4.a Commercial / Institutional" 
# "1.B Fugitive Emissions From Fuels"  
# "2. Industrial Processes" 
# "3. Agriculture"  
# "5. Waste"

#total file of CDNProvGHGs

```

```{r,include=FALSE}
#Import all AU files. Extract list of files from the folder, then use this to parse the data
filenames <- list.files(path = "Data_AUCDN_GHGBreakdown/AGEISbyState", pattern = NULL, all.files = FALSE,
           full.names = FALSE, recursive = FALSE,
           ignore.case = FALSE, include.dirs = FALSE, no.. = FALSE)

AUStateGHGs <- NULL
for (filename in filenames) {
  temp <- read_csv(paste0("Data_AUCDN_GHGBreakdown/AGEISbyState/",filename),n_max=1,col_names=TRUE)
  temp <- select(temp,-X28)
  #parse the filename using regular expressions
  region <- str_extract(filename, "^[A-Z]+")
  sector <- sub("^[A-Z]+[_]{1}","",filename)
  sector <- sub(".csv{1}","",sector)
  temp <- mutate(temp,sector=sector,region=region,cty="AU")
  AUStateGHGs <- rbind(AUStateGHGs,temp)
} 

#gather data
AUStateGHGs <- AUStateGHGs %>% gather("year","GHGs",-sector,-region,-cty)

#generate the sector for merging
Sector_def$sector <- gsub("[[:space:]]{1}.*","",Sector_def$IPCC_Sector)
Sector_def$sector <- gsub("[\\.]$","",Sector_def$sector)

#merge in IPCC definitions using sector
AUStateGHGs <- merge(AUStateGHGs,subset(Sector_def,select=c(sector,IPCC_Sector)),by=c("sector"),all=FALSE)
AUStateGHGs <- select(AUStateGHGs,-sector)
Sector_def <- select(Sector_def,-sector)
#unique(temp$IPCC_Sector)

AUStateGHGs$year <- as.numeric(AUStateGHGs$year)

#need to produce the "other fossil fuel" category for AU data
unique(AUStateGHGs$IPCC_Sector)

# "Total" (to verify totals, but note this includes LULUCF)
# "1.A.1 Energy Industries" (to find 1.A.1.z)
# "1.A.1.a Public Electricity and Heat Production"
# "1.A.1.z Fossil Fuel Extraction" (to create this!)
#   from 1.A.1 minus 1.A.1.a
# "1.A.2 Manufacturing Industries and Construction" 
# "1.A.3 Transport" 
# 1.A.4 Buildings (created from sum of:)
#   "1.A.4.c Agriculture, Forestry and Fishing"
#   "1.A.4.b Residential"
#   "1.A.4.a Commercial / Institutional" 
# "1.B Fugitive Emissions From Fuels"  
# "2. Industrial Processes" 
# "3. Agriculture"  
# "5. Waste"

#simplify dataset:
temp <- AUStateGHGs
AUStateGHGs <- temp
AUStateGHGs <- filter(AUStateGHGs,IPCC_Sector %in% c("Total","1.A.1 Energy Industries","1.A.1.a Public Electricity and Heat Production","1.A.2 Manufacturing Industries and Construction","1.A.3 Transport","1.A.4.c Agriculture, Forestry and Fishing","1.A.4.b Residential","1.A.4.a Commercial / Institutional","1.B Fugitive Emissions From Fuels","2. Industrial Processes","3. Agriculture","5. Waste","3. Agriculture","4. Land Use, Land-Use Change and Forestry KP"))

#How to extract 1.A.1.a from 1.A.1? subset this, take difference? turn to columns, and extract?
  
test <- filter(AUStateGHGs,IPCC_Sector %in% c("1.A.1 Energy Industries","1.A.1.a Public Electricity and Heat Production"))
test <- spread(test,key=IPCC_Sector,value=GHGs)
test$"1.A.1.z Fossil Fuel Extraction and Refining" <- test$"1.A.1 Energy Industries"-test$"1.A.1.a Public Electricity and Heat Production"
test <- gather(test,key = IPCC_Sector,value=GHGs,-year,-cty,-region)
test <- filter(test,IPCC_Sector=="1.A.1.z Fossil Fuel Extraction and Refining")

#Append this to AUStateGHGs
AUStateGHGs <- rbind(AUStateGHGs,test)
rm(test)
unique(AUStateGHGs$IPCC_Sector)
AUStateGHGs <- filter(AUStateGHGs,IPCC_Sector!="1.A.1 Energy Industries")

AUStateGHGs$IPCC_Sector[AUStateGHGs$IPCC_Sector=="1.A.4.a Commercial / Institutional" ] <- "1.A.4 Buildings"
AUStateGHGs$IPCC_Sector[AUStateGHGs$IPCC_Sector=="1.A.4.b Residential"] <- "1.A.4 Buildings"
AUStateGHGs$IPCC_Sector[AUStateGHGs$IPCC_Sector=="1.A.4.c Agriculture, Forestry and Fishing"] <- "1.A.4 Buildings"

AUStateGHGs <- aggregate(GHGs~.,data=AUStateGHGs,FUN=sum)

#verify that total is close?
temp1 <- subset(AUStateGHGs,AUStateGHGs$IPCC_Sector=="Total")
temp2 <- subset(AUStateGHGs,AUStateGHGs$IPCC_Sector!="Total") %>% select(-cty) %>% group_by(region,year) %>% summarise(sectorsum=sum(GHGs))

ggplot() + theme_bw() + facet_grid(cols=vars(region)) +
  geom_line(data=temp1,aes(year,GHGs),color='red',size=5) +
  geom_line(data=temp2,aes(year,sectorsum),color='green')

#totals match well here. Interesting that most of the deforestation is QT

```

```{r}
#Combine AU and Canadian Data
region_IPCC_Data <- rbind(AUStateGHGs,CDNProvGHGs)
region_IPCC_Data <- filter(region_IPCC_Data,IPCC_Sector!="4. Land Use, Land-Use Change and Forestry KP" & IPCC_Sector!="Total")
str(region_IPCC_Data)
unique(region_IPCC_Data$IPCC_Sector)

#merge in convenient names: but first, must generate a new Convenient_Name2 vector?
region_IPCC_Data <- merge(region_IPCC_Data,subset(Sector_def,select=c(IPCC_Sector,Convenient_Name2)),by=c("IPCC_Sector"),all.x=TRUE,all.y=FALSE)
region_IPCC_Data <- rename(region_IPCC_Data,Convenient_Name=Convenient_Name2)

region_IPCC_Data$Convenient_Name[region_IPCC_Data$IPCC_Sector=="1.A.1.z Fossil Fuel Extraction and Refining"] <- "Coal, Oil, and Gas"
region_IPCC_Data$Convenient_Name[region_IPCC_Data$IPCC_Sector=="1.A.4 Buildings"] <- "Buildings"
unique(region_IPCC_Data$Convenient_Name)

#convert to million tonnes (Mt) of CO2
region_IPCC_Data$GHGs <- region_IPCC_Data$GHGs*1000 /(10^6)

#order full dataset
region_IPCC_Data <- region_IPCC_Data[with(region_IPCC_Data,order(cty,region,Convenient_Name,year)),]

#ready to plot
```

```{r,fig.width=12,fig.height=5,include=TRUE}
#plot using facets, by sector? Plot separately for Canada and AU, use common ordering based on total emissions
#try sorting by totals across both countries?

temp <- subset(region_IPCC_Data,select=c(GHGs,Convenient_Name)) %>% group_by(Convenient_Name) %>% summarise(sectorsum=sum(GHGs)) %>% arrange(-sectorsum)
factororder <- as.character(temp$Convenient_Name)

#start with Canada
CDN_region_plot <- region_IPCC_Data %>% filter(cty=="CDN") %>% select(-IPCC_Sector,-cty)

CDN_region_plot$region = factor(CDN_region_plot$region, levels=c('BC','AB','SK','MB','ON','QC','NB','PE','NS','NL'))
CDN_region_plot <- CDN_region_plot[with(CDN_region_plot,order(region,Convenient_Name,year)),]

ggplot(CDN_region_plot,aes(year,GHGs,fill=factor(CDN_region_plot$Convenient_Name,levels=factororder,ordered=TRUE))) + 
  geom_area() + theme_bw() + facet_grid(cols=vars(region)) +
  scale_x_continuous(breaks=seq(1990,2010,20)) + 
  scale_fill_brewer(palette="Spectral") + 
  labs(title="Canadian Greenhouse Gas Emissions", subtitle="Excluding LULUCF",  
       y=expression('Emissions (MT CO'[2]*')'),
       x=element_blank(),
       fill="Category") + geom_col(width=1)

```

```{r,fig.width=12,fig.height=5,include=TRUE}
#Australia 
AU_region_plot <- region_IPCC_Data %>% filter(cty=="AU") %>% select(-IPCC_Sector,-cty)

factororder <- filter(AU_region_plot,AU_region_plot$year==2016 & region=="NSW") %>% arrange(-GHGs)
factororder <- as.character(factororder$Convenient_Name)

#levels <- unique(AU_region_plot$region)
#c('BC','AB','SK','MB','ON','QC','NB','PE','NS','NL')

AU_region_plot$region = factor(AU_region_plot$region, levels=c("NSW","QT","VIC","SA","WA","NT","TAS"))
AU_region_plot <- AU_region_plot[with(AU_region_plot,order(region,Convenient_Name,year)),]

ggplot(AU_region_plot,aes(year,GHGs,fill=factor(AU_region_plot$Convenient_Name,levels=factororder,ordered=TRUE))) + 
  geom_area() + theme_bw() + facet_grid(cols=vars(region)) + scale_fill_brewer(palette="Spectral") + 
  scale_x_continuous(breaks=seq(1990,2010,10)) + 
  labs(title="Australian Greenhouse Gas Emissions", subtitle="Excluding LULUCF", 
       y=expression('Emissions (MT CO'[2]*')'),
       x=element_blank(),
       fill="Category") + geom_col(width=1)
```



`r knitr::knit_exit() #inline r code to end kniting`

```{r}

```


### Appendix: Data details


#### Canada's Emission Targets

  * Toronto target. 
    * See "Peter Usher (1989) World Conference on the Changing Atmosphere:
Implications for Global Security, Environment: Science and Policy for Sustainable Development,
31:1, 25-27, DOI: 10.1080/00139157.1989.9929931" [pdf here](https://www.tandfonline.com/doi/pdf/10.1080/00139157.1989.9929931)
    * *Reduce CO2 emissions by approximately 20% of 1988 levels by the
year 2005 as an initial global goal.*
    * Mention in HCSC ENSU report ["KYOTO AND BEYOND: MEETING THE CLIMATE CHANGE CHALLENGE"](http://www.ourcommons.ca/DocumentViewer/en/36-1/ENSU/report-2/page-72) 
    * Defined in percentage terms only; does not specify LULUCF treatment.
        + Debatable whether this is a government target; the original statements by Mulroney and his environment minister don't appear to be digitized. 
        + Appears in the mycommons ENSU committee report, but the footnote citation is broken
    
*   Rio target
    + Canada signs the UNFCCC at the 1992 Rio Earth Summit. The wording isn't clear, but does appear to contain a commitment by developed countries to stabilize GHGs at 1990 levels by 2000. See this [1992 GoC Summary](http://www.res.parl.gc.ca/Content/LOP/ResearchPublicationsArchive/bp1000/bp317-e.asp#3.%20The%20Convention%20on%20Climate%20Change%20and%20Canada(txt)) 
    + In 1990 Brian Mulroney released the Green Plan which, the internet reports, contained a target of "*Stabilization of carbon dioxide and other greenhouse gas
emissions at 1990 levels by the year 2000.*" 
    + [PDF here from the GoC](http://cfs.nrcan.gc.ca/pubwarehouse/pdfs/24604.pdf)
    + Text from Rio of the UNITED NATIONS FRAMEWORK CONVENTION
ON CLIMATE CHANGE (see DL report) appears to allow for LULUCF in defining stabilization 
* Kyoto Target: [6% below 1990 levels by 2012.](https://unfccc.int/process/the-kyoto-protocol)
    + https://unfccc.int/files/national_reports/initial_reports_under_the_kyoto_protocol/application/pdf/initial_report_of_canada.pdf
    + http://www.ec.gc.ca/doc/ed-es/p_123/CC_Plan_2007_e.pdf
    + includes LULUCF, and defines 1990 emissions (including LULUCF) as 599 MtCO2eq
        + raises question of what "Historical emissions" includes in Canada's trend that I've plotted? 
* Copenhagen Target: [17% reduction from 2005 levels by 2020.](https://www.ec.gc.ca/doc/publications/cc/COM1374/ec-com1374-en-es.htm)
    + https://www.ec.gc.ca/GES-GHG/default.asp?lang=En&n=E0533893-&printfullpage=true&wbdisable=false
    + Defines: "Under the Copenhagen Accord, Canada committed to reducing its emissions by 17% from 2005 levels by 2020.Footnote 1 As economy-wide emissions in 2005 were 736 Mt, Canada’s implied Copenhagen target is 611 Mt in 2020."
* Paris Target: 30% below 2005 levels by 2030.
    + Unclear how sources and sinks will be counted. Canada's NC7 (3rd biennial report) specifies 517 as the target, apparently excluding LULUCF
    + Biennial report 2 specifies a target of 524, which is 30% below the 2005 level of 749. This excludes LULUCF

### New data process

Need to reconcile a) whether LULUCF is included in the emissions, b) differences between how emissions for various targets were measured and c) why my numbers differ from those of the carbon action tracker. 

MORE:
* Toronto target
    + Defined in percentage terms only; does not specify LULUCF treatment.
    + Debatable whether this is a government target; the original statements by Mulroney and his environment minister don't appear to be digitized. 
        + Appears in the mycommons ENSU committee report, but the footnote citation is broken
        
        
# Data & Details


Comparing greenhouse gas emissions between Canada and Australia is, surprisingly to me, not straight forward. While both countries release emissions data (thanks, unnamed workers!) they don't do so in a way that's at all convenient. I'll leave the data deets for the end, in case you're interested. 

## GHG Data Details

The UNFCCC specifies countries report their GHG emissions through [National Inventory Reports](https://unfccc.int/process/transparency-and-reporting/reporting-and-review-under-the-convention/greenhouse-gas-inventories-annex-i-parties/national-inventory-submissions-2018)  using a common reporting framework. This framework allows some comparability between countries as the [IPCC lists reporting requirements](https://unfccc.int/process-and-meetings/transparency-and-reporting/reporting-and-review-under-the-convention/greenhouse-gas-inventories-annex-i-parties/reporting-requirements) for GHG inventory submissions. While this lists the sectors, it does not define them.

There are two complications to comparing AU and CDN data. First, Canada does not appear to provide public emissions data using the IPCC CRF definitions specified by the IPCC. Second, Australia does not provide data using sector definitions that are easily readable by software. (there are multiple categories called "other", distinguishable only by context) As such, I construct a basic cross-walk from both AU and CDN data to IPCC NIR CRF tables definitions, and then from CRF definitions to convenient sector names that closely matches Canada's sector definitions.

This [German website](https://iir-de.wikidot.com/1-a-1-a-public-electricity-and-heat-production) has some definitions of the sectors, though I'm not sure where they came from. 

### Canada

#### Canadian GHG Data Sources

The Government of Canada provides [a list of emission data and reports](https://www.canada.ca/en/environment-climate-change/services/climate-change/greenhouse-gas-emissions.html)

*  These include the link to Canada's greenhouse gas inventory [Canada's greenhouse gas inventory](https://www.canada.ca/en/environment-climate-change/services/climate-change/greenhouse-gas-emissions/inventory.html)
    + This page links to [Greenhouse gas data and other information](https://open.canada.ca/data/en/dataset/779c7bcf-4982-47eb-af1b-a33618a05e5b) which in turn links to the [ECCC Datamart](http://data.ec.gc.ca/data/substances/monitor/national-and-provincial-territorial-greenhouse-gas-emission-tables/)
  * Data in the ECCC DataMart matches the NIR report, and contains Canada's historical GHG trends.
  * The *National Inventory Report 1990–2015: Greenhouse Gas Sources and Sinks in Canada, Environment and Climate Change Canada, 2017* links to [UNFCCC submissions on GHGS.](https://unfccc.int/process/transparency-and-reporting/reporting-and-review-under-the-convention/greenhouse-gas-inventories-annex-i-parties/national-inventory-submissions-2018)
    + Part 3 of the NIR lists tables of the emissions data, and states electronic versions are available at http://www.open.canada.ca
  * *Canada’s Seventh National Communication on Climate Change and Third Biennial Report—Actions to meet commitments under the United Nations Framework Convention on Climate Change, Environment and Climate Change Canada, 2017*[PDF](https://unfccc.int/files/national_reports/national_communications_and_biennial_reports/application/pdf/82051493_canada-nc7-br3-1-5108_eccc_can7thncomm3rdbi-report_en_04_web.pdf) 
    + Data in the report that can be used to cross-check with other emissions data.
* Auditor General Report [Perspectives on Climate Change Action in Canada — A Collaborative Report from Auditors General](http://www.oag-bvg.gc.ca/internet/English/parl_otp_201803_e_42883.html)
    + This lists projected emissions on page 35/44.
* Emissions from the Auditor General report don't appear to mach the NIR emissions
  + This may be because the NIR 2018 updates past emissions. For example:
    + Auditor General 2016 projections report 722Mt. This report uses 2017 NIR tables
    + NIR 2018 tables for 2016 emissions IPCC categories reports 704 Mt
    + NIR 2017 tables for 2016 emissions Canadian Economic Sector reports 722
    + NIR 2018 tables for 2016 emissions Canadian Economic Sector reports 704

### Australia

#### Australia GHG Data Sources

Australia provides only a patchwork of GHG data, summarized on the [Department of Environment and Energy](http://www.environment.gov.au/climate-change/climate-science-data/greenhouse-gas-measurement/) page.

  * This links to the a variety of datasets, including [the National Inventory, Quarterly Updates, State, and Sector data.](http://www.environment.gov.au/climate-change/climate-science-data/greenhouse-gas-measurement/publications#national)
  * None of these has a full dataset on AU emissions from 1990 to now. Individual years or state data can be accessed via AGEIS, but not a table of all data. 
  * Using AGEIS [link here](http://ageis.climatechange.gov.au/NGGI.aspx), apparently CIF/IPCC Sector categories can be downloaded.
    * Data must be saved by individual year. Sector categories are not unique; for example, there are multiple "other" categories.
* [Australia emission projections are available here.](http://www.environment.gov.au/climate-change/climate-science-data/emissions-projections) and [here](http://www.environment.gov.au/climate-change/publications/emissions-projections-2017)
    * These include an excel files with 1990-2030 emissions and projections, with projections presumably starting in 2017 though this isn't clear. 
  * Using AGEIS, downloading state-industry-year emissions by Kyoto classification.
  


